PAMELA: an annotation-based Java Modeling Framework

### 2021-04-22 ResponseLetterMinor.tex

\bigskip
Sincerely yours,

\bigskip
Sylvain, Guillaume, Caine, Joel, Jean-Christophe, Salvador, Fabien and Antoine

\pagebreak

\Reviewer{\#1}
\begin{response}{There remain or are introduced however, a substantial number of typos, language errors, and other minor issues that need correcting:}

All highlighted typos, languages errors and minor issues have been checked and fixed.

%\begin{itemize}

%\item "to /avoid" on p3 (DONE)
%\item "detailed thereafter" $->$ "detailed hereafter" (It's Correct : As adverbs the difference between hereafter and thereafter is that hereafter is in time to come; in some future time or state while thereafter is after that, from then on.)
%\item "increasing PAMELA metamodel" $->$ "by \_extending\_ PAMELA\_'s\_ metamodel" (DONE)
%\item "with setter/getter": make both plural (DONE)
%\item POJO used here and elsewhere yet the acronym is introduced much later (DONE)
%\item p4 l11-12: reduce etc. should be reduce\_s\_ etc. (Keep as it is)
%\item odd spacing before : consistently throughout bullet list + text  (Keep as it is: trust Latex)
%\item "in the section 3.2" $->$ "in Section 3.2" (style: named references deserve a capital and no article before) (DONE)
%\item p5 and many times: "PAMELA metamodel", "PAMELA model", "PAMELA framework" need "the" before them. (Done ; with Pamela $->$ PAMELA)
%\item "defines API": insert "an" (DONE)
%\item "in Java interface": insert "a" (DONE)
%\item "returning desired value", "taking value" etc.: insert "the" (DONE)
%\item "May also declares"... (DONE)
%\item p6 l46: @Embedded/"as embedded" meaning?
%\item p7 "of launched Java application" missing "the" (DONE)
%\item "is/are" $->$ "is (are)", in line with parentheses earlier in sentence (DONE)
%\item "Experimented programmers": I doubt it, unless you got major ethical clearance (Replaced by "Experienced PAMELA programmers")
%\item "Listing 2: model" $->$ "~: Model" consistent with Listing 1 (DONE)
%\item the book title is "Lord of the ring\_s\_" with s (DONE)
%\item "using classical Java ... scheme": missing "the" (DONE)
%\item "Java programming language": insert "The" (DONE)
%\item l43 inconsistent spacing (DONE)
%\item l46 package ... package (DONE)
%\item p11 l36 sticking into margin (Final check to do)
%\item S.4.3 please add size in LoC and classes (?)
%\item java $->$ Java (DONE)
%\item p13 "diana component": insert "the" (DONE)
%\item "experimentation" $->$ "experiment" (DONE)
%\item "When Pamela has been" $->$ "After Pamela was" (DONE : if we consider it's over)
%\item l36 the quotation mark before code generation is the wrong one 'changed to ' and ' ?)
%\item p 14 "appertaining" $->$ "pertaining", "relating" (Kept as is. As verbs the difference between pertain and appertain is that pertain is to belong while appertain is to belong to or be a part of, whether by right, nature, appointment, or custom; to relate to.)
%\end{itemize}

\end{response}

%\pagebreak

\Reviewer{\#2}
\begin{response}{Would it be possible to structure it more, because currently it is several paragraphs the one bellow the others.}

Section 2 has been decomposed into 2 subsections

\end{response}

\begin{response}{figure 2 is not enough described in section 2 and even in the paper.}

The end of 2.1 subsection provides an explanation of figure 2.

\end{response}

\begin{response}{More over, concerning section 4.4, I appreciate this new section. However, while reading, it is not a good advertisement for Pamela. Overhead (x20) is huge. Perhaps it is not the right benchmark if it provides bad results that are not representing the right usage.}

SM comments

\end{response}

### 2021-04-22 ResponseLetterMinor.tex

\end{response}

%\pagebreak

\Reviewer{\#2}
\begin{response}{Would it be possible to structure it more, because currently it is several paragraphs the one bellow the others.}

Section 2 has been decomposed into two Subsections: Subsection 2.1.  Architecture overview
and Subsection 2.2.  Usage of PAMELA.

@@

\end{response}

\begin{response}{figure 2 is not enough described in section 2 and even in the paper.}

The end of Subsection 2.1 provides now a full explanation of figure 2.

@@

\end{response}

\begin{response}{More over, concerning section 4.4, I appreciate this new section. However, while reading, it is not a good advertisement for Pamela. Overhead (x20) is huge. Perhaps it is not the right benchmark if it provides bad results that are not representing the right usage.}

We agree with the reviewer that the overhead may seem huge. Note, however, that these measures correspond to the very worst-case scenario: an application with absolutely no business logic. Although such applications do not correspond to the normal usage, we believe showing this measurement is important to fully characterize the performance of the approach. E.g., PAMELA should not be used when the application is a mere data structure.

For normal applications, the time passed on attribute accessors is negligible compared to the time spent on business logic. We have verified this by using a profiling tool on PAMELA based applications (we have added this information to the paper). Moreover, as we show in Section 5 PAMELA has been successfully applied to many industrial projects, with no performance issues reported.

### 2021-04-22 implementation.tex
small corrections completed. Remain: improve fig2 comments, restructure section 2, rephrase 4.4
Adding minor revision response letter
restructuration section 2
typos in response letter 2

The \texttt{org.openflexo.pamela} package contains the following feature-related Java interfaces:
\begin{itemize}
    \item \texttt{AccessibleProxyObject} is the interface that PAMELA objects should extend in order to benefit from base features such as generic default implementation, containment management, notification, object graph comparison and diff/merge, visiting patterns, etc.
    \item \texttt{CloneableProxyObject} exposes features related to cloning.
    \item \texttt{DeletableProxyObject} exposes features related to deletion management.
    \item \texttt{SpecifiableProxyObject} exposes dynamic assertion checking features in the context of JML (contract management) use. 
\end{itemize}

@@

\begin{itemize}
    \vspace{-0.2cm}\item Core implementation (\texttt{pamela-core}) provides all base features for PAMELA. It contains 20k lines of code involving 184 classes. This code is covered with unit tests (6k lines of code, 112 classes). Reached code coverage is about 66\%.
    \vspace{-0.2cm}\item \texttt{pamela-security-patterns} is an add-on library containing some security-pattern implementations.
    \vspace{-0.2cm}\item \texttt{pamela-perf-tests} contains benchmarking tools whose purpose is to quantify PAMELA performances. 
\end{itemize}

@@

\subsection{On performance issues}

Compared to a basic POJO implementation, use of PAMELA involves a significant CPU and memory-footprint overhead, because of the partially-interpreted nature for some model features.

A workbench\footnote {\url{https://github.com/openflexo-team/pamela/tree/1.6.1/pamela-perf-tests}} has been developed to help measure that overhead. We use a base model composed of four entities defining 4 to 5 properties (single and multiple properties), and we instantiate that tree model (1010101 objects created and initialized to default value). From this simple model, two Java implementations have been derived using code generation. The first implementation was made using fully implemented Java classes (POJO), while the second one is fully interpreted (it uses the PAMELA framework and defines only interfaces and API methods).

Unsurprisingly, we quantify a significant CPU overhead (from $\times$20). This is due to the fact that \texttt{MethodHandler} mechanism is fully interpreted and contains many hooks, and cannot be compared to a fully compiled and optimized Java dynamic binding. As well, memory footprint is increased from an 
$\times$5 factor.

These conclusions must be moderated by the fact that this overhead only applies
to a very small portion of the code (mostly property accessors methods). A "real-world" model implementation generally involves a bigger "business code" part, which is generally the first CPU-time consumer. Using PAMELA won’t increase CPU use on "business code" if this code is implemented with plain Java code (which is generally the case). As we will see in the next section, PAMELA has been successfully applied to many industrial projects, where no major performance issues were reported.

Moreover, comparison between the two implementations should be tempered by the fact that the PAMELA implementation offers many functional features which are not present in the base implementation (such as undo/redo, clipboard management, runtime monitoring and weaving, etc.).

>>>>

The \texttt{org.openflexo.pamela} package contains the following feature-related Java interfaces:
\begin{itemize}
    \item \texttt{AccessibleProxyObject} is the interface that PAMELA objects should extend in order to benefit from base features such as generic default implementation, containment management, notification, object graph comparison and diff/merge, visiting patterns, etc.
    \item \texttt{CloneableProxyObject} exposes features related to cloning.
    \item \texttt{DeletableProxyObject} exposes features related to deletion management.
    \item \texttt{SpecifiableProxyObject} exposes dynamic assertion checking features in the context of JML (contract management) use. 
\end{itemize}
% Joel : tu fais toujours référence à JML mais tu pourrais juste mentionner le Contract management ou Design By Contract approach ??

@@

\begin{itemize}
    \vspace{-0.2cm}\item Core implementation (\texttt{pamela-core}) provides all base features. %for PAMELA. 
    It contains 20k lines of code involving 184 classes. This code is covered with unit tests (6k lines of code, 112 classes). Reached code coverage is about 66\%.
    \vspace{-0.2cm}\item \texttt{pamela-security-patterns} is an add-on library containing some security-pattern implementations.
    \vspace{-0.2cm}\item \texttt{pamela-perf-tests} contains benchmarking tools whose purpose is to quantify PAMELA performances. 
\end{itemize}

@@

\subsection{Performance analysis}

Compared to a basic POJO implementation, the use of PAMELA implies a CPU and memory-footprint overhead. This is  due to the partially-interpreted nature of some of its features. We have developed a performance workbench in order to measure it \footnote {\url{https://github.com/openflexo-team/pamela/tree/1.6.1/pamela-perf-tests}}.

We first defined a base model composed of four entities declaring four to five properties (single and multi-valued) each. From this simple model, two Java implementations have been derived using code generation. The first implementation uses fully implemented Java classes (POJO), while the second one is fully interpreted (it uses the PAMELA framework and defines only interfaces and API methods). Then, for each implementation, we have instantiated the base model (by creating 1010101 objects with their properties initialized to default values) and carried out performance measures.  As expected, we measured a CPU overhead for the fully interpreted implementation (from $\times$20). This is due to \texttt{MethodHandler} mechanism which is fully interpreted and contains many hooks. Therefore, it  cannot be compared to a fully compiled and optimized Java dynamic binding. We also measured a memory footprint overhead, but of a less important weight that the CPU overhead ($\times$5 factor).

Note however that the measured overhead only applies to a very small portion of the code (mostly property accessors methods).  A "real-world" model implementation generally involves a bigger "business code" part, which is generally the first CPU-time consumer. Using PAMELA won’t increase CPU use on "business code" if this code is implemented with plain Java code (which is generally the case). We have used the Yourkit Java Profiler\footnote{https://www.yourkit.com/} on PAMELA based applications and, as expected, the time passed on accessors was negligible compared to the time used on methods implementing business logic. As we will see in the next section, PAMELA has been successfully applied to many industrial projects, with no performance issues reported. Moreover, the PAMELA implementation offers many functional features which are not present in the base implementation (such as undo/redo, clipboard management, runtime monitoring and weaving, etc.) which out-weight the performance overhead.
